---
title: "final project"
execute:
  keep-md: true
  df-print: paged
  warning: false
  echo: false
format:
  html:
    code-fold: true
    code-line-numbers: true
    fig-width: 10
date: "`r format(Sys.time(), '%B %d, %Y')`"
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(readxl)
library(readr)
library(dplyr)
library(rnaturalearth)
library(sf)
```

```{r}
wbb_rosters23 <- read.csv("https://github.com/Sports-Roster-Data/womens-college-basketball/raw/main/wbb_rosters_2022_23.csv")
```

More than 25 teams have at least seven non-U.S. players on their rosters, South Florida has 12, San Francisco 10 and six teams have nine each (Simon Fraser has 14 but it is located in Canada).

```{r}
# More than 25 teams have at least seven non-U.S. players on their rosters, South Florida has 12, San Francisco 10 and six teams have nine each (Simon Fraser has 14 but it is located in Canada).
wbb_rosters23 %>% 
  filter(country_clean != 'USA') %>% 
  group_by(team) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))
```

Of the 95 teams that have at least one Australian player, 25 of them have two or more on their roster. Southern Nazarene WBB has *six* Aussies.

```{r}
# Of the 95 teams that have at least one Australian player, 25 of them have two or more on their roster. Southern Nazarene WBB has _six_ Aussies.

wbb_rosters23 %>% 
  filter(country_clean == 'AUSTRALIA') %>% 
  group_by(team) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))

wbb_rosters23 %>% 
  filter(country_clean == 'SPAIN') %>% 
  group_by(team) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))
```

```{r}
total_players <- nrow(wbb_rosters23)
```

```{r}
proportions <- wbb_rosters23 %>%
  summarise(
    International = sum(!is.na(country_clean) & country_clean != 'USA') / total_players,
    US = sum(country_clean == 'USA') / total_players
  )
```

```{r}
proportions_long <- proportions %>%
  gather(key = "PlayerType", value = "Proportion")
```

Looking at the proposition of international and U.S players

```{r}
ggplot(proportions_long, aes(x = "", y = Proportion, fill = PlayerType)) +
  geom_bar(stat = "identity") +
  coord_polar("y") +
  labs(title = "Proportion of International and U.S. Players",
       fill = "Player Type") +
  theme_minimal() +
  theme(legend.position = "right",
        axis.text = element_blank(),  # Remove x-axis and y-axis labels
        axis.title = element_blank())+
   geom_text(aes(label = scales::percent(Proportion)), position = position_stack(vjust = 0.5), size = 4)
```

```{r}
international_data <- wbb_rosters23 %>%
  filter(!is.na(position_clean) & !is.na(country_clean))%>%
  filter(country_clean != 'USA') %>%
  group_by(position_clean, country_clean) %>%
  summarize(position_count = n(),
            avg_height = mean(total_inches)) %>%
  arrange(desc(position_count))
```

Looking at the total of international in each position. Found out Guard and Center are the two common positions from other countries.

```{r}
ggplot(international_data, aes(x = position_clean, y = position_count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Total International Players by Position",
       x = "Position",
       y = "Count") +
  theme_minimal() 
```

Looking at the average height between International players and US players

```{r}
# Filter out rows with missing position, total_inches, or country values
filtered_data <- wbb_rosters23 %>%
  filter(!is.na(position_clean) & !is.na(total_inches) & !is.na(country_clean))

# Calculate the average total_inches for each position and country
average_heights <- filtered_data %>%
  na.omit(position_clean)%>%
  filter(position_clean != 'GUARD/CENTER')%>%
  group_by(position_clean, country_clean) %>%
  summarize(avg_total_inches = mean(total_inches, na.rm = TRUE),
            position_count = n())

# Combine international and U.S. player data
combined_data <- average_heights %>%
  mutate(country_clean = ifelse(country_clean == 'USA', 'US', 'International'))

# Create a bar chart
ggplot(combined_data, aes(x = position_clean, y = avg_total_inches, fill = country_clean)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Height of Players by Position",
       x = "Position",
       y = "Average heights(inches)",
       fill = "Country") +
  theme_grey() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

are certain positions or player types more common among international players compared to U.S. players?

```{r}
total_position_counts <- combined_data %>%
  group_by(position_clean, country_clean) %>%
  summarize(total_count = sum(position_count))
```

Looking for which positions are common in US and International

```{r}
# Create a bar chart
ggplot(total_position_counts, aes(x = position_clean, y = total_count, fill = country_clean)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of International and U.S. Players by Position",
       x = "Position",
       y = "Counts",
       fill = "Position Count") +
  theme_grey() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}

# Filter for guards and centers
international_guards_centers <- international_data %>%
  filter(position_clean == c("GUARD", "FORWARD"))%>%
  slice_max(order_by = position_count, n = 3)
```

```{r}
ggplot(international_guards_centers, aes(x = position_clean, y = position_count, fill = country_clean)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of International Players by Position",
       x = "Position",
       y = "Counts",
       fill = "Position Count") +
  theme_grey()
```
