---
title: "W6: AboutTime"
execute:
  keep-md: true
  df-print: paged
  warning: false
  echo: false
format:
  html:
    code-fold: true
    code-line-numbers: true
    fig-width: 10
date: "`r format(Sys.time(), '%B %d, %Y')`"
editor_options: 
  chunk_output_type: console
---

```{r}
library(priceR)
library(Lahman)
library(tidyverse)
library(ggrepel)
library(knitr)

#price_data <- tq_get(c("BA", "AMZN", "MSFT"),
#       get = "stock.prices",
#       from = "2020-03-01")
#You can add "to = " also to further refine the dates
```

```{r}
data <- read.csv(url("https://byuistats.github.io/M335/data/sales.csv")) %>% 
  mutate(Time = lubridate::ymd_hms(Time, tz = "UTC")) %>% 
  mutate(Time = with_tz(Time, "America/Denver"))

df_count <- data %>% 
  group_by(Amount) %>% 
  summarise(count = n())
```

```{r}
data2<- data %>% filter(Name != "Missing")%>%
  group_by(Name, Month = floor_date(Time, "month")) %>% 
  summarise(Amount = sum(Amount)) %>% 
  ggplot(aes(x = Month, y = Amount, color = Name)) +
  geom_line() +
  labs(title = "Monthly Gross Revenue", x ="Date", y = "Gross Revenue") +
  theme_bw()

data2
```

```{r}
data1<- data %>% filter(Name != "Missing")%>%
  group_by(Name, Month = floor_date(Time, "month")) %>% 
  summarise(Transactions = n()) %>% 
  ggplot(aes(x = Month, y = Transactions, color = Name)) +
  geom_line() +
  labs(title = "Monthly Customer Traffic", x ="Month", y = "Number of Transactions") +
  theme_bw()

data1
```

```{r}
data3 <- data %>%
  filter(Name != "Missing") %>%
  mutate(Hour = lubridate::hour(Time)) %>%
  group_by(Name, Hour) %>%
  summarise(Transactions = n())

ggplot(data3, aes(x = Hour, y = Transactions, fill = Name)) +
  geom_area() +
  labs(
    title = "Hourly Customer Traffic",
    x = "Hour of the Day",
    y = "Number of Transactions",
    fill = "Company"
  ) +
  theme_bw()
```

Based on the visualizations and data analysis, it appears that HotDiggity has performed the best over the three-month period. HotDiggity consistently maintained a higher level of gross revenue and customer traffic compared to other companies. The hourly customer traffic analysis also reveals that HotDiggity has a more extended and active operating window throughout the day, indicating a strong presence in the market. This, coupled with their consistent performance, suggests that HotDiggity has effectively capitalized on customer transactions and generated higher revenue. However, it's essential to conduct a more in-depth analysis to understand the specific factors contributing to their success and make a comprehensive assessment.
