---
title: "W5: BYU Baseball"
execute:
  keep-md: true
  df-print: paged
  warning: false
  echo: false
format:
  html:
    code-fold: true
    code-line-numbers: true
    fig-width: 10
date: "`r format(Sys.time(), '%B %d, %Y')`"
editor_options: 
  chunk_output_type: console
---

```{r}
library(Lahman)
library(priceR)
library(tidyverse)
library(pander)
```

```{r}
dat <- Schools %>% 
  full_join(CollegePlaying) %>% 
  filter(state == "UT") %>% 
  select(name_full, playerID, yearID) %>% 
  group_by(playerID) %>% 
  mutate(most_recent_year = max(yearID)) %>% 
  distinct(name_full, playerID) %>% 
  left_join(People, by = "playerID") %>% 
  unite("full_name", nameGiven, nameLast, sep = " ") %>% 
  select(playerID, name_full, full_name) %>% 
  left_join(Salaries, by = "playerID") %>% 
  select(-teamID) %>% 
  rename("school" = name_full, "player" = full_name) %>% 
  # drop all rows with N/A
  na.omit()

# create a new column to adjust inflation 
dat$salary_2020 <- 
  adjust_for_inflation(price = dat$salary, 
                       from_date = dat$yearID, 
                       country = "US", 
                       to_date = 2020)
```

```{r}
dat <- dat %>% 
  select(-playerID)

dat1 <- dat %>% 
  head()

pander(dat1)
```

```{r}
byu_dat <- dat %>% 
  group_by(school) %>% 
  summarise(ave_salary = mean(salary_2020))

require(scales)
ggplot(byu_dat)+
  geom_bar(aes(y = fct_reorder(school, ave_salary, .desc = TRUE), x = ave_salary, fill = school), stat="identity")+
    theme_bw()+
  # labels = comma remove the scientific annotation 
  scale_x_continuous(labels=scales::dollar_format())+
  theme(axis.text.x = element_text(angle = -45, hjust =0))+
  xlab(NULL)+
  ylab(NULL)+
  labs(
    title = "BYU Players Performance Compared to Other Schools Based on Average Salaries ",
    subtitle = "The average salary for BYU alumini players is $3195489.1"
  )+
  theme(legend.position = "none")
```

```{r}
dat %>% 
  group_by(playerID) %>% 
  ggplot(aes(x = yearID, y = salary_2020, color = player, shape = school))+
  geom_point(size = 3)+
  annotate("text", x = 2009, y = 8745970, label = "Jeremy Guthrie\n from BYU")+
  annotate("text", x = 1989, y = 8900000, label = "John Scott Morris\n from BYU")+
  annotate("text", x = 1999, y = 7800000, label = "Wallace Kith Joyner\n from BYU")+
  scale_y_continuous(labels=scales::dollar_format())+
  theme_classic()+
  ylab(NULL)+
  xlab(NULL)+
  labs(
    title = "Scatter Plot of Individual Player's Salary\nWho Graduated from University in Utah ", 
    subtitle = "Most players are BYU alumini, including top 3 successful players"
  )+
  theme(
    legend.position = "bottom", 
    legend.title = element_blank()
  )+
  guides(color=FALSE)+
  geom_vline(xintercept = 1990, color = "grey")+
  geom_vline(xintercept = 2000, color = "grey")+
  geom_vline(xintercept = 2010, color = "grey")+
  geom_hline(yintercept = 7500000, color = "red", linetype ="dashed")
```

## Summarize

From chart 1, we compared the average salaries earned by players graduating from universities in Utah. We found out that BYU alumni players have more success than other Utah schools alumni. As for chart 2, the scatter plot supports our finding that BYU alumni players are more successful. Three of players who got paid the most all graduated from Brigham Young University. Moreover, it is shown that most players who graduated from Utah school are from BYU.
